Indirect elicitation as a necessary condition for consistent surrogate losses

### 2020-06-02 neurips-20/cvx-flats.tex
Minor tweaks on wording and addressing comments/deleting FYIs

While indirect elicitation is only a necessary condition, our bounds recover or improve upon the state of the art:
we achieve tighter bounds than the feasible subspace dimension approach for discrete predictions~\citep{ramaswamy2016convex}\raft{correct cite?\jessie{yep}}, yet recover their result (Corollary~\ref{cor:fsd-bound}), and no bounds for convex surrogates were known in the continuous prediction setting~\citep{frongillo2018elicitation}.
Furthermore, in all of the examples we consider (\S~\raf{examples}), our lower bounds match the best upper bounds.
For these reasons, we advocate for indirect elicitation as a tractable and more general alternative to calibration when studying the consistency of surrogates.

@@

% After reviewing previous work and tools to study consistency (\S~\ref{sec:related-work}), we lay down a new general definition of consistency which can simultaneously capture all four settings mentioned above (\S~\ref{sec:consis-implies-indir}).
% Our main result (\S~\ref{sec:char-convex}, Theorem~\ref{thm:cvx-flats}) provides a necessary condition for
% the existence of consistent convex surrogates in a given dimension $d$.
% This result is closely related to those provided by~\citet[Theorem 16]{ramaswamy2016convex} in the discrete prediction setting; in fact, we show their result is a special case of our main result (Corollary~\ref{cor:fsd-bound}).
% We illustrate our main result with several examples (\S~\raf{examples}), one of which addresses an open question posed by~\citet[Section 8]{frongillo2018elicitation} in the continuous estimation setting.
\raft{FYI: use citet when treating citations as nouns; citep is used when the citation does not participate in the grammar, like: ``That is funny [Noah Alexander Frongillo, personal communication, 5/31/2020].''  Not sure what 'cite' does in this style...}\jessiet{Noted, and propogated}

@@

We consider supervised learning problems in the space $\X \times \Y$, for some \emph{feature space} $\X$ and \emph{label space} $\Y$m with data drawn from a distribution $D$ over $\X \times \Y$.
The task is to produce a hypothesis $f: \X \to \R$, for some \emph{prediction space} $\R$, which may be different from $\Y$.
For example, in ranking problems $\R$ may be all $n!$ permutations over the $n$ labels forming $\Y$.
As we focus heavily on conditional distributions $D_x = \Pr[Y|X=x]$ over $\Y$ given some $x \in X$, we abstract away $x$ with the following notation.
The probability simplex over $\Y$ is $\simplex$.
\raft{FYI: moved $\supp$ to where we use it.\jessie{ok}}

@@

Arising from the statistics and economics literature, property elicitation is similar to calibration, but only places a restriction on the exact minimizers of a surrogate~\citep{savage1971elicitation,osband1985information-eliciting,lambert2008eliciting,lambert2009eliciting,lambert2018elicitation,frongillo2015vector-valued,frongillo2014general}.
Specifically, given a statistic or \emph{property} $\Gamma$ of interest, which maps a distribution $p$ over $\Y$ to the set of desired or correct predictions, the minimizers of $L$ should precisely coincide with $\Gamma$.
Intuitively, $p = \Pr[Y|X=x]$ is a conditional distribution, though the definition is also applied to point prediction settings.
\raft{FYI Jessie: I like this line but $D_x$ has not been defined yet \jessie{I added $D$ in the notation section and introduced it there.  Let me know if it really doesn't mesh with you, but I think it fit somewhat naturally.}}

@@

It is straightforward to show that, for a discrete target loss $\ell$, calibration implies indirect elicitation, as perhaps first observed by \citet{agarwal2015consistent}.
Formally, if $(L,\psi)$ is calibrated with respect to $\ell$, then $(L,\psi)$ indirectly elicits $\gamma = \prop{\ell}$.
To see this, take $u\in\U$ and $p\in\Gamma_u$, implying $u\in\Gamma(p)$.
From eq.~\eqref{eq:elic}, $\exploss{L}{u}{p} = \inf_{u'\in\U} \exploss{L}{u'}{p}$, so we must have $\psi(u) \in \gamma(p)$ from eq.~\eqref{eq:calibration}, as desired.
\raft{Jessie: if this argument makes sense to you, you can cut that prop in the appendix now}
As a result, indirect elicitation in a necessary condition for consistency when given a target discrete loss.
As we will see in \S~\ref{sec:consis-implies-indir}, it is also a necessary condition for consistency more broadly, including in all four quadrants mentioned above.
\btw{Some refs / related work commented out here.  I think not as relevant as elicitation complexity.}

@@

\subsection{Elicitation complexity notions}\label{subsec:complexity}

Various works have studied the minimum prediction dimension $d$ needed in order to construct a consistent surrogate loss $L: \reals^d \times \Y \to \reals$, typically through proxies such as calibration~\citep{ramaswamy2016convex} and property elicitation~\citep{frongillo2015vector-valued,fissler2015higher,frongillo2018elicitation}.
\raft{Nice!  Love ``proxies'' here.}
In the case of a discrete target loss, \citet{ramaswamy2016convex} introduce \emph{convex calibration dimension}, the minimum prediction dimension yielding a convex calibrated surrogate.
Their results has led to the design of consistent convex surrogates for discrete prediction problems such as hierarchical classification~\citep{ramaswamy2015hierarchical} and classification with an abstain option~\citep{ramaswamy2018consistent}.
\raft{FYI Jessie: Two general points.  (1) I spent the majority of my time in this section just on dependencies; e.g.\ you had ``similar to elicitation complexity mentioned above'' but elicitation complexity was not mentioned above.  Next time it would be more efficient if you made sure to balance local edits with global scans to catch these sorts of things, as well as making sure we aren't explaining things twice, etc.  (2) You tend to mix background with nods to our results, which is actually good to do, but it can be jarring if you go back and forth too much.  See for example how I changed the previous subsection, reserving the nod to our work to the end of the subsection.}

@@

In related work,
\citet{finocchiaro2019embedding} study surrogate losses which \emph{embed} a discrete loss in some sense, which is a special case of indirect elicitation.
They construct a calibrated link functions from embeddings, which together with observations about indirect elicitation, imply that for polyhedral (piecewise linear and convex) losses, calibration and indirect elicitation are equivalent.
\citet{finocchiaro2020embedding} further introduce the notion of \emph{embedding dimension}, which is a lower bound on both convex elicitation complexity of discrete properties and convex calibration dimension of discrete losses.
Since embedding is a special case of indirect elicitation, we have the chain of inequalities, embedding dimension $\leq$  ``polyhedral calibration dimension'' $\leq$ convex calibration dimension, but it is an open problem whether any of these inequalities can be strict.
\raft{FYI Jessie: nice job!  Tightened it up a bit.}

@@

\botodo{I felt the reader needed a bit more transition here, and added one. Also I think this proof sketch was much too dense for all but the most expert in convex analysis! I slowed things down, and see next note.}
Thanks to Theorem \ref{thm:consistent-implies-indir-elic}, the task is now this: given a target property $\gamma$, or loss $\ell$ with $\gamma = \prop{\ell}$, prove lower bounds on any $d$ such that some convex $L: \reals^d \times \Y \to \reals$ indirectly elicits $\gamma$.
We now explore a tool, Theorem \ref{thm:cvx-flats}, for proving such convex-elicitation lower bounds.
The key idea, crystallized from the proof of \citet[Theorem 16]{ramaswamy2016convex}, is to consider a particular optimal surrogate report $u \in \reals^d$.
\botodo{wording/credit ok? \jessie{I think so}}
Theorem \ref{thm:cvx-flats} will show that if $d$ is small, then the level set $\{p : u \in \argmin_{u'} \exploss{L}{u'}{p}\}$ \bo{used comma instead of semicolon} must be large; in fact, it must roughly contain a high-dimensional affine subspace, or \emph{flat}.
But by definition of indirect elicitation, there is some level set $\gamma_r$ (where $u$ is linked to $r$) containing this flat as well.
The point of this result is the contrapositive: if $\gamma$ has a more intricate level set that does not contain any high-dimensional flats, then $\gamma$ cannot have a low-dimensional consistent surrogate.

\botodo{Not sure the proof sketch is worth it, since the proof itself is pretty short. Commenting out for now, but let me know what y'all would prefer!}
\botodo{TODO: we have not yet introduced subgradients or the subdifferential!}\jessiet{Added as a footnote since I would hope most people know what a subgradient is, but also a good point to not assume.}
%As a first step, observe that all minima of convex functions are global, so $\vec 0  \in \partial L(u, y)$ if and only if $u$ minimizes $L$ in its first component.
%This allows us to consider subgradient sets of the loss at a fixed distribution $p$ as the weighted Minkowski sums of subgradient sets for the loss on each outcome.\bo{This won't be clear except to readers very well versed in convex analysis. I think we'll have to sketch why that's true.}
%Fixing a distribution $p \in \simplex$, we construct a set of $n$ witnesses in $\partial L(u,y)$ for each $y$ whose $p$-weighted sum is $\vec 0$.
%The nullspace of the matrix of these witnesses yields a flat, formally introduced below, of codimension at most $d$ intersecting our fixed distribution $p$ and the level set(s) $\gamma_r$ containing $p$.

\raf{Introduce $\affhull$}

@@

\section{Omitted Proofs}\label{app:omitted-proofs}
\begin{proposition}\label{prop:elic-cplx-lb-ccdim}
	Consider the discrete loss $\ell : \R \times \Y \to \reals_+$ and $\gamma:= \prop{\ell}$.
	Then $\eliccvx(\gamma) \leq \ccdim(\ell)$.
\end{proposition}
\begin{proof}
	Let $d$ be the convex calibration dimension of $\ell$, and $(L, \psi)$ be calibrated with respect to $\ell$ with $L$ convex.
	Then $L$ indirectly elicits $\gamma$ by Lemma~\ref{lem:calib-implies-indir}.
\end{proof}
\subsection{Reconstructing the result of~\citet[Theorem 16]{ramaswamy2016convex}}
\begin{lemma}\label{lem:feas-sub-is-a-flat}
	Suppose we have the discrete elicitable property $\gamma$ and distribution $p \in \relint{\simplex}$ with $r \in \gamma(p)$.
	If $F$ is a flat in $\affhull(\simplex)$ containing $p$ such that $F \cap \simplex \subseteq \gamma_r$, then $F - p$ is a subspace contained in $\Sc_{\gamma_r}(p)$.
	%  \raf{I think you want $p$ in the interior of the simplex for now}
\end{lemma}
\begin{proof}
	Observe $F-p$ is a subspace as it is a linear shift of $F$, which is a linear subspace by definition of a flat and the fact that it contains $\vec 0$.
	Now consider $v \in F - p$.
	Since $p \in \relint{\simplex}$, there is an open ball of radius $\epsilon$ in the affine hull of $\simplex$ so that for all $q \in B(p, \epsilon)$, we have $q \in \simplex$.
	In particular, take $\alpha = \epsilon / 2$, and we observe $p \pm \alpha v \in B(p, \epsilon)$, and therefore $p \pm \alpha v \in \simplex$.
	Moreover, by the assumption $v \in F - p$, we also have $p \pm \alpha v \in \gamma_r$. 
	Since level sets of elicitable properties are convex (\citep{lambert2009eliciting}) this is true for all $\alpha' \leq \alpha$.
	Therefore, we observe $v \in S_{\gamma_r}(p)$, so $F-p \subseteq S_{\gamma_r}(p)$.
	%  First, if $p+v \in \gamma_r$ and $p -v \in \gamma_r$, then we have $v \in \Sc_{\gamma_r}(p)$ with $\epsilon_0 = 1$ as level sets of elicitable properties are convex by~\cite{lambert2009eliciting}.
	%  If either $p + v$ or $p - v \not \in \gamma_r$, it must be because the term is out of the simplex by definition of $F$.
	%  
	%  However, if both $p + \alpha^+ v$ and $p - \alpha^- v \in \simplex$ for some $\alpha^\pm \in (0,1)$, then $v \in \Sc_{\gamma_r}(p)$ with $\epsilon_0 = \min(\alpha^+, \alpha^-)$.
	%  As $p \in \relint{\simplex}$, there is always such an $\epsilon_0$; if there were not, then we would observe $v \not \in F - p$.
	%  Therefore, we have $v \in F - p \implies v \in \Sc_{\gamma_r}(p)$, so $(F - p) \subseteq \Sc_{\gamma_r}(p)$.
\end{proof}

>>>>

While indirect elicitation is only a necessary condition, our bounds recover or improve upon the state of the art:
we achieve tighter bounds than the feasible subspace dimension approach for discrete predictions~\citep{ramaswamy2016convex}, yet recover their result (Corollary~\ref{cor:fsd-bound}), and no bounds for convex surrogates were known in the continuous prediction setting~\citep{frongillo2018elicitation}.
Furthermore, in all of the examples we consider, our lower bounds match the best upper bounds.
For these reasons, we advocate for indirect elicitation as a tractable and more general alternative to calibration when studying the consistency of surrogates.

@@

% After reviewing previous work and tools to study consistency (\S~\ref{sec:related-work}), we lay down a new general definition of consistency which can simultaneously capture all four settings mentioned above (\S~\ref{sec:consis-implies-indir}).
% Our main result (\S~\ref{sec:char-convex}, Theorem~\ref{thm:cvx-flats}) provides a necessary condition for
% the existence of consistent convex surrogates in a given dimension $d$.
% This result is closely related to those provided by~\citet[Theorem 16]{ramaswamy2016convex} in the discrete prediction setting; in fact, we show their result is a special case of our main result (Corollary~\ref{cor:fsd-bound}).
% We illustrate our main result with several examples (\S~\raf{examples}), one of which addresses an open question posed by~\citet[Section 8]{frongillo2018elicitation} in the continuous estimation setting.
%\raft{FYI: use citet when treating citations as nouns; citep is used when the citation does not participate in the grammar, like: ``That is funny [Noah Alexander Frongillo, personal communication, 5/31/2020].''  Not sure what 'cite' does in this style...}\jessiet{Noted, and propogated}

@@

We consider supervised learning problems in the space $\X \times \Y$, for some \emph{feature space} $\X$ and \emph{label space} $\Y$m with data drawn from a distribution $D$ over $\X \times \Y$.
The task is to produce a hypothesis $f: \X \to \R$, for some \emph{prediction space} $\R$, which may be different from $\Y$.
For example, in ranking problems $\R$ may be all $n!$ permutations over the $n$ labels forming $\Y$.
As we focus heavily on conditional distributions $D_x = \Pr[Y|X=x]$ over $\Y$ given some $x \in X$, we abstract away $x$ with the following notation.
The probability simplex over $\Y$ is $\simplex$.

@@

Arising from the statistics and economics literature, property elicitation is similar to calibration, but only places a restriction on the exact minimizers of a surrogate~\citep{savage1971elicitation,osband1985information-eliciting,lambert2008eliciting,lambert2009eliciting,lambert2018elicitation,frongillo2015vector-valued,frongillo2014general}.
Specifically, given a statistic or \emph{property} $\Gamma$ of interest, which maps a distribution $p$ over $\Y$ to the set of desired or correct predictions, the minimizers of $L$ should precisely coincide with $\Gamma$.
Intuitively, $p = \Pr[Y|X=x]$ is a conditional distribution, though the definition is also applied to point prediction settings.
%\raft{FYI Jessie: I like this line but $D_x$ has not been defined yet \jessie{I added $D$ in the notation section and introduced it there.  Let me know if it really doesn't mesh with you, but I think it fit somewhat naturally.}}

@@

It is straightforward to show that, for a discrete target loss $\ell$, calibration implies indirect elicitation, as perhaps first observed by \citet{agarwal2015consistent}.
Formally, if $(L,\psi)$ is calibrated with respect to $\ell$, then $(L,\psi)$ indirectly elicits $\gamma = \prop{\ell}$.
To see this, take $u\in\U$ and $p\in\Gamma_u$, implying $u\in\Gamma(p)$.
From eq.~\eqref{eq:elic}, $\exploss{L}{u}{p} = \inf_{u'\in\U} \exploss{L}{u'}{p}$, so we must have $\psi(u) \in \gamma(p)$ from eq.~\eqref{eq:calibration}, as desired.
%\raft{Jessie: if this argument makes sense to you, you can cut that prop in the appendix now}
As a result, indirect elicitation in a necessary condition for consistency when given a target discrete loss.
As we will see in \S~\ref{sec:consis-implies-indir}, it is also a necessary condition for consistency more broadly, including in all four quadrants mentioned above.
\btw{Some refs / related work commented out here.  I think not as relevant as elicitation complexity.}

@@

\subsection{Elicitation complexity notions}\label{subsec:complexity}
\jessiet{Suggest: Notions of prediction dimension, or something like it.}

Various works have studied the minimum prediction dimension $d$ needed in order to construct a consistent surrogate loss $L: \reals^d \times \Y \to \reals$, typically through proxies such as calibration~\citep{ramaswamy2016convex} and property elicitation~\citep{frongillo2015vector-valued,fissler2015higher,frongillo2018elicitation}.
%\raft{Nice!  Love ``proxies'' here.}
In the case of a discrete target loss, \citet{ramaswamy2016convex} introduce \emph{convex calibration dimension}, the minimum prediction dimension yielding a convex calibrated surrogate.
Their results has led to the design of consistent convex surrogates for discrete prediction problems such as hierarchical classification~\citep{ramaswamy2015hierarchical} and classification with an abstain option~\citep{ramaswamy2018consistent}.
%\raft{FYI Jessie: Two general points.  (1) I spent the majority of my time in this section just on dependencies; e.g.\ you had ``similar to elicitation complexity mentioned above'' but elicitation complexity was not mentioned above.  Next time it would be more efficient if you made sure to balance local edits with global scans to catch these sorts of things, as well as making sure we aren't explaining things twice, etc.  (2) You tend to mix background with nods to our results, which is actually good to do, but it can be jarring if you go back and forth too much.  See for example how I changed the previous subsection, reserving the nod to our work to the end of the subsection.}\jessiet{Noted boss man}

@@

In related work,
\citet{finocchiaro2019embedding} study surrogate losses which \emph{embed} a discrete loss in some sense, which is a special case of indirect elicitation.
They construct a calibrated link functions from embeddings, which together with observations about indirect elicitation, imply that for polyhedral (piecewise linear and convex) losses, calibration and indirect elicitation are equivalent.
\citet{finocchiaro2020embedding} further introduce the notion of \emph{embedding dimension}, which is a lower bound on both convex elicitation complexity of discrete properties and convex calibration dimension of discrete losses.
Since embedding is a special case of indirect elicitation, we have the chain of inequalities, embedding dimension $\leq$  ``polyhedral calibration dimension'' $\leq$ convex calibration dimension, but it is an open problem whether any of these inequalities can be strict.
%\raft{FYI Jessie: nice job!  Tightened it up a bit.}

@@

\botodo{I felt the reader needed a bit more transition here, and added one. Also I think this proof sketch was much too dense for all but the most expert in convex analysis! I slowed things down, and see next note.}
Thanks to Theorem \ref{thm:consistent-implies-indir-elic}, the task is now this: given a target property $\gamma$, or loss $\ell$ with $\gamma = \prop{\ell}$, prove lower bounds on any $d$ such that some convex $L: \reals^d \times \Y \to \reals$ indirectly elicits $\gamma$.
We now explore a tool, Theorem \ref{thm:cvx-flats}, for proving such convex-elicitation lower bounds.
The key idea, crystallized from the proof of \citet[Theorem 16]{ramaswamy2016convex}, is to consider a particular optimal surrogate report $u \in \reals^d$.
Theorem \ref{thm:cvx-flats} will show that if $d$ is small, then the level set $\{p : u \in \argmin_{u'} \exploss{L}{u'}{p}\}$ must be large; in fact, it must roughly contain a high-dimensional affine subspace, or \emph{flat}.
But by definition of indirect elicitation, there is some level set $\gamma_r$ (where $u$ is linked to $r$) containing this flat as well.
The point of this result is the contrapositive: if $\gamma$ has a more intricate level set that does not contain any high-dimensional flats, then $\gamma$ cannot have a low-dimensional consistent surrogate.

\botodo{Not sure the proof sketch is worth it, since the proof itself is pretty short. Commenting out for now, but let me know what y'all would prefer!}
%As a first step, observe that all minima of convex functions are global, so $\vec 0  \in \partial L(u, y)$ if and only if $u$ minimizes $L$ in its first component.
%This allows us to consider subgradient sets of the loss at a fixed distribution $p$ as the weighted Minkowski sums of subgradient sets for the loss on each outcome.\bo{This won't be clear except to readers very well versed in convex analysis. I think we'll have to sketch why that's true.}
%Fixing a distribution $p \in \simplex$, we construct a set of $n$ witnesses in $\partial L(u,y)$ for each $y$ whose $p$-weighted sum is $\vec 0$.
%The nullspace of the matrix of these witnesses yields a flat, formally introduced below, of codimension at most $d$ intersecting our fixed distribution $p$ and the level set(s) $\gamma_r$ containing $p$.

\raf{Introduce $\affhull$}
Throughout, we often use the affine hull of the simplex, $\affhull(\simplex) = \{\sum_{i=1}^n \alpha_i p_i : p_i \in \simplex, \alpha_i \in \reals, \sum_{i=1}^n \alpha_i = 1\}$, which has dimension $n-1$.

@@

\section{Omitted Proofs}\label{app:omitted-proofs}

\subsection{Reconstructing the result of~\citet[Theorem 16]{ramaswamy2016convex}}
\begin{lemma}\label{lem:feas-sub-is-a-flat}
	Suppose we have the discrete elicitable property $\gamma$ and distribution $p \in \relint{\simplex}$ with $r \in \gamma(p)$.
	If $F$ is a flat in $\affhull(\simplex)$ containing $p$ such that $F \cap \simplex \subseteq \gamma_r$, then $F - p$ is a subspace contained in $\Sc_{\gamma_r}(p)$.
	%  \raf{I think you want $p$ in the interior of the simplex for now}
\end{lemma}
\begin{proof}
	Observe $F-p$ is a subspace as it is a linear shift of $F$, which is a linear subspace by definition of a flat and the fact that it contains $\vec 0$.
	Now consider $v \in F - p$.
	Since $p \in \relint{\simplex}$, there is an open ball of radius $\epsilon$ in the affine hull of $\simplex$ so that for all $q \in B(p, \epsilon)$, we have $q \in \simplex$.
	In particular, take $\alpha = \epsilon / 2$, and we observe $p \pm \alpha v \in B(p, \epsilon)$, and therefore $p \pm \alpha v \in \simplex$.
	Moreover, by the assumption $v \in F - p$, we also have $p \pm \alpha v \in \gamma_r$. 
	Since level sets of elicitable properties are convex (\citep{lambert2009eliciting}) this is true for all $\alpha' \leq \alpha$.
	Therefore, we observe $v \in S_{\gamma_r}(p)$, so $F-p \subseteq S_{\gamma_r}(p)$.
	%  First, if $p+v \in \gamma_r$ and $p -v \in \gamma_r$, then we have $v \in \Sc_{\gamma_r}(p)$ with $\epsilon_0 = 1$ as level sets of elicitable properties are convex by~\cite{lambert2009eliciting}.
	%  If either $p + v$ or $p - v \not \in \gamma_r$, it must be because the term is out of the simplex by definition of $F$.
	%  
	%  However, if both $p + \alpha^+ v$ and $p - \alpha^- v \in \simplex$ for some $\alpha^\pm \in (0,1)$, then $v \in \Sc_{\gamma_r}(p)$ with $\epsilon_0 = \min(\alpha^+, \alpha^-)$.
	%  As $p \in \relint{\simplex}$, there is always such an $\epsilon_0$; if there were not, then we would observe $v \not \in F - p$.
	%  Therefore, we have $v \in F - p \implies v \in \Sc_{\gamma_r}(p)$, so $(F - p) \subseteq \Sc_{\gamma_r}(p)$.
\end{proof}
